\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-26 coq-example.tex
Reply some of the comments. Fixed Closing Fact part.

Our emphasis here is that, we singled out some language features from
STLC and prove the type safety for each feature separately, following
the style of the examples in MTC~\cite{delaware2013,forsta2020}.
And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature
to mix the semantics and properties of two language features---product
and boolean (in \cref{fig:STLC-example})---with vanilla STLC. This way,
we can say precisely, \textit{a programming language feature itself is a
piece of data/inheritance judgement/inherited family}. Compared with MTC,
our example uses small-step operational semantics\YZ{Does MTC not define a small-step operational semantics for STLC?} by exploiting
extensible inductive types, and most of the proof is directly adapted
from the one in Software Foundation, resulting in a more accessible proof.\YZ{Can you say more about how your plugin enables more code reuse and modularity?}
% In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach. Their advantage is the transferability of this technique accross different proof assistants, and their disadvantage is that their approach are less accessible and unfriendly to amateur Coq users---which can be reflected from the distance of their approach and text-book Software Foundation proof.

% We have problem on inversion lemma. Check if it is the same problem as MTC
We use "Closing Fact" to state and prove the inversion lemma instead of
using the extensible proving mechanism "FTheorem". The reasons are that
(1) it introduces much less boilerplate code because the proof for these
inversion lemmas should be just simple case analysis and we should rely
on Coq to auto-generate them;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?} (2) it shouldn't bother us in the future
because any extension on the syntax should still satisfy these inversion
lemmas; (3) most importantly, we believe this inversion ``lemma'' should
be part of the definition of the syntax instead of considering the
syntax as a mere concrete inductive type. We should postulate this
inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify
that our inductive definition does satisfy the constraint, which is
exactly what we expect from "Closing Fact".

@@

We construct four families, in one inheritance chain.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}

The first family Lang define the big step operator (using fuel) of a while language, with expression. But it postulates the computational information of State used by Lang. It takes about 200+ LOCs.

The second family LangwAbs define the abstract interpreter for it, together with a bunch of postulates, to prove the correctness of the abstract interpreter. It takes about 300+ LOCs.

The third Family LangMore extend LangwAbs with nat constant and addition, and if-then-else control flow, of course retaining all the soundness theorem. It takes about 200+ LOCs.

The fourth family LangRun instantiate some of the postulates of the LangMore and recover computational information of Lang. The concrete interpreter is doing as expected; the abstract interpreter is expected to act as a type-analyzer. It takes about 200+ LOCs.

And at the end, we compile LangRun into module and use Coq to run it on some simple queries.

>>>>

Our emphasis here is that, we singled out some language features from
STLC and prove the type safety for each feature separately, following
the style of the examples in MTC~\cite{delaware2013,forsta2020}.
And we use a \textit{not-yet-formally-defined} \textbf{mix-in} feature
to mix the semantics and properties of two language features---product
and boolean (in \cref{fig:STLC-example})---with vanilla STLC. This way,
we can say precisely, \textit{a programming language feature itself is a
piece of data/inheritance judgement/inherited family}. Compared with MTC,
our example uses small-step operational semantics\YZ{Does MTC not define a small-step operational semantics for STLC?}\EDJreply{I thought they use big step with fuel. Though Coq a la carte use small-step in one example} by exploiting
extensible inductive types, and most of the proof is directly adapted
from the one in Software Foundation, resulting in a more accessible proof.\YZ{Can you say more about how your plugin enables more code reuse and modularity?}
% In related work, Coq/Metatheory a la carte/Tion embedding can be emphasized as a more "semantic approach" because they encode the meaning using a special design pattern (for example, open recursive inductive type for extension), compared to our more syntactic approach. Their advantage is the transferability of this technique accross different proof assistants, and their disadvantage is that their approach are less accessible and unfriendly to amateur Coq users---which can be reflected from the distance of their approach and text-book Software Foundation proof.

% We have problem on inversion lemma. Check if it is the same problem as MTC
We use "Closing Fact" to state and prove the inversion lemma instead of
using the extensible proving mechanism "FTheorem". The reasons are that
(1) it introduces much less boilerplate code because the proof for these
inversion lemmas should be just simple case analysis and we should rely
on Coq to generate most of the boilerplate code;\YZ{I was under the impression that the plugin could not auto-generate Closing Facts or their proofs, no?}\EDJreply{I have make the above Closing Fact section more about this detail. Please check.} (2) it shouldn't bother us in the future
because any extension on the syntax should still satisfy these inversion
lemmas; (3) most importantly, we believe this inversion ``lemma'' should
be part of the definition of the syntax instead of considering the
syntax as a mere concrete inductive type. We should postulate this
inversion ``lemma'' like \textit{a constraint} and post-hoc-ly verify
that our inductive definition does satisfy the constraint, which is
exactly what we expect from "Closing Fact".

@@

We construct four families, in one inheritance chain.\YZ{Why is it useful to define these concrete and abstract interpreters as an inheritance chain?}\EDJreply{I don't think these examples are here to show the example themselves are useful. They are here to show family inheritance is useful, and can extend stuff to all sorts of things. I also add some details below on how each family can be extended and what each family means.}

The first family "Lang" define the big step operator (using fuel) of a while language, with expression. But it postulates the computational information of "State" used by Lang. It takes about 200+ LOCs. We can swap the implementation of "State" to have different memory/state management by extending family "Lang".

The second family "LangwAbs" define the abstract interpreter for it, together with a bunch of postulates, to prove the correctness of the abstract interpreter. It takes about 300+ LOCs. This example shows how to use family inheritance to extend a concrete interpreter with abstract interpreters. This example also shows in this family polymoprhism framework we can reason about "Lang" with computation details being abstracted.

The third Family "LangMore" extend LangwAbs with nat constant and addition, and if-then-else control flow, of course retaining all the soundness theorem. It takes about 200+ LOCs. This is one example of using family inheritance to support new language feature, and compatible with the existent reasoning.

The fourth family "LangRun" instantiate some of the postulates of the LangMore and recover computational information of Lang. The concrete interpreter is doing as expected; the abstract interpreter is expected to act as a type-analyzer. It takes about 200+ LOCs. This is one example of instantiate the detail computation. In this example, we use a simple Coq function to simulate state-mapping information.

And at the end, we compile LangRun into module and use Coq to run it on some simple queries. This "LangRun" also illustrates that we still have computability in the presence of family.

### 2022-06-26 coqimpl.tex
Reply some of the comments. Fixed Closing Fact part.

% This "Closing Fact" also can ease us from the necessity to let plugin
% generate proof for recursors and partial recursors and their
% computational rules (as in \ref{langdesign-req4}), and thus making
% plugin development easier.
\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy. Question is if it makes the programmer's life easier.}\EDJreply{You are right. It does sound like a bad excuse. I rewrite the whole section now. Closing Fact is initially used when the user wants global reasoning instead of extensible proof by induction, for example, the inversion lemma. \\ It just happens that they can be used by me so that I can postpone generation of these rule plugin (which is hard for me to do). This paragraph is just me being honest that my plugin is not mature enough to generate everything.}
% For example, the computation rule for a
% concrete recursor like "B.neg" and partial recursors in
% \cref{fig:plugin-example2} can be directly asserted by using "Closing
% Fact" and their proofs verified later during compilation.
\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{(1) For partial recursor and recursor, that should be generated by plugin (2) For general global reasoning, I don't know. There must be some other ways in the literature. }
% This ``Global Reasoning'' will be again useful in later examples.\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}

@@

Thus to refer to them, we have to scope the defining tactic with "MetaData" and our plugin will setup the appropriate environment. After a "MetaData" block is closed, the plugin will aggregate the content into the current defining family so that they can also be used when defining the following fields of the family. This can fulfill \ref{langdesign-req5}.

To fulfill \ref{langdesign-req5}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of

@@

% To remedy this, we support a "MetaData" command. This command
% bundles any original Coq primitive (in the current family
% context) into the compiled module type, and makes them visible to the
% following defining field. Doing so, we can define customized tactic as
% in \ref{langdesign-req5} that uses lemmas proved in the family.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}

>>>>

% This "Closing Fact" also can ease us from the necessity to let plugin
% generate proof for recursors and partial recursors and their
% computational rules (as in \ref{langdesign-req4}), and thus making
% plugin development easier.
\YZ{This reads like a bad excuse: the programmer does not care about if the plugin development is easy. Question is if it makes the programmer's life easier.}\EDJreply{You are right. It does sound like a bad excuse. I rewrite the whole section now. Closing Fact is initially used when the user wants global reasoning instead of extensible proof by induction, for example, the inversion lemma. \\ It just happens that they can be used by me so that I can postpone generation of these rule plugin (which is hard for me to do). This paragraph is just me being honest that my plugin is not mature enough to generate everything.}
% For example, the computation rule for a
% concrete recursor like "B.neg" and partial recursors in
% \cref{fig:plugin-example2} can be directly asserted by using "Closing
% Fact" and their proofs verified later during compilation.
\YZ{What is the alternative if not using "Closing Fact"?}\EDJreply{(1) For partial recursor and recursor, that should be generated by plugin (2) For general global reasoning, I don't know. There must be some other ways in the literature. }\YZ{Still not entirely clear to me what is gained and what is lost by using "Closing Fact".}\EDJreply{Please check.}
% This ``Global Reasoning'' will be again useful in later examples.

@@

Thus to refer to them, we have to scope the defining tactic with "MetaData" and our plugin will setup the appropriate environment. After a "MetaData" block is closed, the plugin will aggregate the content into the current defining family so that they can also be used when defining the following fields of the family. This can fulfill \ref{langdesign-req5}.\YZ{Consider adding examples of Closing Fact and MetaData, even if only to be included in an appendix.}

% To fulfill \ref{langdesign-req5}, we need somehow to aggregate tactic inside compiled module type, due to the implementation of our family---when defining every field of

@@

% To remedy this, we support a "MetaData" command. This command
% bundles any original Coq primitive (in the current family
% context) into the compiled module type, and makes them visible to the
% following defining field. Doing so, we can define customized tactic as
% in \ref{langdesign-req5} that uses lemmas proved in the family.
